\OurTool{}: Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles

### 2020-06-29 Response-NAR-Round1.tex

Over the course of this extensive round of revisions, we have:
\begin{itemize}
  \item Rewritten extensively our manuscript, notably including the bulk of our method section;
  \item Assessed the significance of our main quantitative claims as per \cite{Xu2012}, and included a new paragraph describing the methodology;
  \item Performed additional tests to ascertain the quality of reactivities produced for \didy: We analyzed the correlations of reactivity profiles at the reactivity level (Supp. Sec. 3.2); We also found significantly better agreement, with the reference structure of \cite{Meyer2014}, of all produced reactivity profiles than expected at random (Supp. Sec. 3.1);
  \item Improved the distribution, documentation, and streamlined the installation of \OurTool (although we could not reach a \Software{bioconda} distribution yet);
  \item Fixed numerous typos and unfortunate turns of sentences.
\end{itemize}

We hope that, upon reviews, the editor and reviewers will share our enthusiasm for \OurTool, and consent to its dissemination to the wide readership of Nucleic Acids Research.

@@

			Moreover, we observe that conditions clustered from the Ensemble Distance metrics (due to their impact of the pseudo-Boltzmann ensemble) do not always correlate in term of raw reactivity, with very poor correlations occurring within clustered conditions. This essentially reflects the fact that similar profiles may induce very similar pseudo-Boltzmann ensembles, \emph{ie} different premises sometimes imply the same conclusions.

			Conversely, one observes very few strong correlations outside of clusters. In fact, the only two exceptions are \OneMSevILU and \OneMSevILUThree which, despite belonging to different clusters and achieving different qualities of prediction, correlate almost perfectly with each other and with other reactivity profiles produced using 1M7/SHAPE-MaP. Other natural candidates for explaining the reactivity correlations (Map vs CE, +Mg vs -Mg) do not appear to be supported by this analysis.

			Overall, this analysis confirms the consistency of the produced probing profiles (positive correlation), but also suggests a very delicate effect of reactivity data on the downstream predictions, with very different profiles leading to similar predictions (\OneMSevILU), and similar profiles inducing very different ones (\OneMSevILU vs \OneMSevILUThree).}

				In particular, while reactivity profiles correlate positively in a pairwise fashion, as can be expected, these results do not support the hypothesis that the method of detection (CE vs Map) would represent a strong differentiator at the level of reactivity profiles. In fact, the only clear clusters are:
		\begin{itemize}
			\item  $\{\NMIAMg,\NMIA\}$;
			\item  $\{\NMIACE,\NAICE,\OneMSevCE\}$;
			\item $\{\OneMSevILU, \OneMSevILUThree,\OneMSevILUThreeMg,\OneMSevILUMg\}$,
		\end{itemize} having respectively a similar reagent, technology, or both, as a common trait. However, neither of those determinants seem universally supported by the observation of the correlation matrix.

@@

	\Comment{Much space is allocated to outlier detection (cluster B) and then to describing analyses with and without cluster B. I feel that the authors can safely designate cluster B as an outlier and exclude it from the analyses described in the main text. Analyses with cluster B can be described in the Supp.}

	\Answer{Again, we respectfully disagree with this vision. What is important about cluster B is not only the fact that it can be identified from our new metrics (more conclusively than than from the raw reactivities), but also that it can be used to assess the impact of contamination by an erroneous/outlier condition.}

	\Comment{Were all the analyzed datasets normalized by any standard method (e.g., 2\%-8\% or box-plot)? I noticed that the mutate-and-map dataset was renormalized, and to the best of my knowledge, the Hajdin dataset is also normalized. However, I didn’t see any information re: Cordero dataset, and more importantly re: the newly generated data analyzed here. Normalization is especially important in the context of differential analysis between conditions.}

	\Answer{
	All reactivities were (re)normalized using the boxplot of \cite{Deigan2009}.

	It is important to clarify that our analysis does not rely on a direct differential analysis at the reactivity level, except in our newly added Supplementary Section 3.2. The latter features a cross-analysis based on the Pearson correlation, precisely chosen as its results are impervious to preprocessing consisting of an affine transforms of the data.
}

@@

		This number may not seem to be supported by a inspection Figure 5 (PCA, left above), but it must be reminded that a PCA is simply a projection. Points that are distant in the PCA are typically distant in the original higher-dimensional space, but close points in the PCA may also be quite distant due to the missing depth perception.

		On Figure 5, we already see a clear separation between B, D, C, A, and H. Moreover, merging clusters E, F and G would lead to the aggregation of substantially different conditions, as can be seen in Figure 6 (right above), so we considered 8 conditions.}

	\Comment{The authors compare stop-based to mutation-based experiments. However, the term “stop-based” is somewhat misleading because there is another variable here - the modification assay platform. MaP protocols must be assayed by NGS whereas the stop-based protocols used in this work rely on CE for modification detection. It is possible that the differences observed between MaP and CE data (see e.g., Fig. 5) have little to do with the stop strategy and more with the sequencing platform (CE vs NGS). The use of “stop-based” might create an impression that this feature is necessarily responsible for these differences. The ideal comparison should be between experiments such as SHAPE-MaP and SHAPE-seq, but given that experiments were already performed, I recommend changing the term and clarifying the technical differences between these subsets of conditions.}

	\Answer{We agree that SHAPE CE should not be mischaracterized as a representative for all SHAPE protocols based on RT stops (notoriously including SHAPE-Seq). Our descriptions were sometimes misleading in that respect, so we modified and removed the corresponding column in Table 2 and tried to remove any sentence that would involuntarily suggest the superiority of a technology/reagent based on our observations.}

@@

	\Comment{Does IPANEMAP perform better when sampling deeper? More generally, was sample size of 1,000 selected following performance optimization or simply because there is a convention in the field that 1,000 suffices for reproducibility at the base-pair level?}

	\Answer{
		We were happy with the overall reproducibility of our results with 1~000 structures, which has been the standard practice in the field since the original \cite{Ding2003} paper. Moreover, our computational requirements scale quadratically with the number of structure, so we recommended a sample size of 1~000, which represents in our opinion a good tradeoff between reproducibility and efficiency.

		To address this question more precisely, we reexecuted the \didy benchmark using a sample size of 2~000, and observed a minor non-significant variation in the MCC (\emph{i.e.} the results did not allow to refute the hypothesis of equal expected MCCs for both experiments).}
